\subsection{Depth-Proportional Resolution (DPR) Policy Algorithm}
\label{sec:depth-proportional-resolution-algo}

The depth-proportional resolution policy algorithm provides capped extant record size with even coverage over record history.
This requires retained observations to be spaced with a gap width proportional to record depth.
%An equivalent interpretation of DPR strategy would be interspersal of the historical record with a fixed number of waypoints.
Alternatively, DPR can be seen as interspersing the historical record with a fixed number of waypoints.

Because observation time points are immutable after the fact, translating this naive DPR plan to a rolling, ``online'' basis necessitates a further consideration.
To conservatively maintain resolution guarantees, it is acceptable to err on the side of caution by choosing gap sizes smaller than the worst-case requirement.
This approach allows a simple trick for achieving policy self-consistency: flooring gap sizes to the next lower power of two.
Under this scheme, gap size will periodically double.
Beacause multiples of a binary power superset multiples of higher binary power, self-consistency is maintained.
As intuition, therefore, the full DPR policy algorithm can be conceived of through a simple principle: %successive eliminations of every other retained observation when a capacity threshold is reached.
each time a capacity threshold is reached, every second observation is eliminated.

% Binary floor approximation exhibits another useful property: policy algorithm property distortions can be bounded by a constant.
% Minimum gap width is half of the highest allowed gap width, while maximum extant record count is twice the lowest allowed count.
Policy algorithm behavior is parametrized by a minimum number of bin windows over record history, $r$.
All gap sizes are equal (or halved), so absolute resolution guarantee of at least $n/r$, with $n$ as record depth, applies.
Further, because binary flooring at most halves gap width, it at most doubles record count.
This property gives the record size cap of $2r$.

Algorithm \ref{alg:depth-proportional-resolution-algo-enum-retained-ranks} provides enumeration of retained time points under the DPR scheme.
Although this process can be achieved via set subtraction between enumerations at successive time points with $\mathcal{O}(1)$ complexity, Algorithm \ref{alg:depth-proportional-resolution-algo-gen-drop-ranks} provides a more expedient approach.
% Theorems \ref{thm:depth-proportional-resolution-algo-uncertainty-bound} and \ref{thm:depth-proportional-resolution-algo-space-complexity} cover extant record size order of growth and resolution guarantorship, respectively.
Figure \ref{fig:retention-policies} includes a time-lapse of the extant record under the DPR policy algorithm.

For simplicity, we have presented a bare-bones approach to depth-proportional resolution, where the entire record is simultaneously decimated by a factor of two upon reaching capacity.
This procedure results in regular episodes where extant record count instantaneously halves.
Such fluctuation may be undesirable.
Many use-cases for constant space complexity will arise from fixed memory allocation.
Such reserved memory cannot typically be used for other purposes, so any unused space would be wasted.
%Occupying only half of bounded capacity clearly comes out on the ``use it or lose it'' downside.

An alternate ``tapered'' variant of the depth-proportional resolution algorithm remedies this space-usage quirk.
%Instead of all at once,
The tapered approach eliminates phased-out observations one by one as new observations accrue, but otherwise has the same properties as the algorithms described for DPR.
% Supplementary Section \ref{sec:depth-proportional-resolution-tapered-algo} details this tapered DPR policy algorithm.
% TODO: ADD SECTION ^^^
The accompanying \texttt{hstrat} software library implements both variants.

% \input{alg/depth-proportional-resolution-algo-pred-keep-rank.tex}.
\input{alg/depth-proportional-resolution-algo-enum-retained-ranks.tex}
\input{alg/depth-proportional-resolution-algo-gen-drop-ranks.tex}
% \input{thm/depth-proportional-resolution-algo-space-complexity.tex}
% \input{thm/depth-proportional-resolution-algo-uncertainty-bound.tex}
